---
title: "Analyse the number of taxa"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyse the number of taxa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Loading this package

```{r}
library(Cer2016)
options(warn = 2)
```


Loading the number of taxa:

```{r}
df_n_taxa <- read_collected_n_taxa()
```

Loading the parameters:

```{r}
df_parameters <- read_collected_parameters()
```


## The success rate

```{r fig.width = 7, fig.height = 7}
n_fail <- sum( is.na(df_n_taxa$n_taxa))
n_success <- sum(!is.na(df_n_taxa$n_taxa))
df_success <- data.frame(
  name = c("success", "fail"), n = c(n_success, n_fail)
)

ggplot2::ggplot(
  df_success, 
  ggplot2::aes(
    x = factor(""), 
    y = df_success$n, 
    fill = df_success$name
  )
) + 
  ggplot2::geom_bar(stat = "identity") + 
  ggplot2::scale_x_discrete("") +
  ggplot2::scale_y_continuous("number of parameter sets") +
  ggplot2::coord_polar(theta = "y") + 
  ggplot2::ggtitle("Incipient species trees created")
```

## What is the global distribution of number of taxa?

Plot the distribution of number of taxa (in the reconstructed incipient species tree) over all runs:

```{r fig.width = 7, fig.height = 7}
df <- df_n_taxa[ !is.na(df_n_taxa$n_taxa), ]

n_missing <- sum(is.na(df_n_taxa$n_taxa))
ggplot2::ggplot(
  data = df, ggplot2::aes(df$n_taxa)
) + 
ggplot2::geom_histogram(bins = 30) + 
ggplot2::ggtitle(
  paste0(
    "Number of taxa for ", n_success, "/",
    nrow(df_n_taxa) ," runs"
  )
)
```


## Which parameters?

Now we create a data frame of both parameters and number of
taxa:

```{r}
# Create a 'filename' column needed for merge
df_parameters$filename <- rownames(df_parameters) 
# Outer join on the filename
df <- merge(df_parameters, df_n_taxa, by = "filename")
# Either remove the NA's...
df <- df[ !is.na(df_n_taxa$n_taxa), ]
# Or keep the failed ones
df_fail <- df[ is.na(df_n_taxa$n_taxa), ]
```

Here we plot both successes and fails. The latter are only plot if present.

```{r fig.width = 7, fig.height = 7}
if (nrow(df_fail) == 0) {
  ggplot2::ggplot(data = df, ggplot2::aes(x = df$sirg, y = df$n_taxa, color = as.factor(df$erg))) + 
  ggplot2::geom_point(shape = 16) +
  ggplot2::ggtitle("Effect of speciation and extinction rates on #taxa") 
} else {
  ggplot2::ggplot(data = df, ggplot2::aes(x = df$sirg, y = df$n_taxa, color = as.factor(df$erg))) + 
  ggplot2::geom_point(shape = 16) +
  ggplot2::geom_jitter(data = df_fail, shape = 3, width = 0.02, height = 0.02, ggplot2::aes(x = df_fail$sirg, y = 0,   color = as.factor(df_fail$erg))) + 
  ggplot2::ggtitle("Effect of speciation and extinction rates on #taxa (crosses denote NA)") 
}
```
