---
title: "Analyse gamma's"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyse gamma's}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Loading this package

```{r}
library(Cer2016)
```

## Collecting gamma's from one or more phylogenies

```{r}
phylogenies <- c(ape::rcoal(10), ape::rcoal(20))
df <- collect_gamma_statistics(phylogenies)
knitr::kable(df)
```

## Collecting gamma's from the species trees

```{r}
filename <- find_path("toy_example_3.RDa")
df <- collect_species_tree_gammas(filename)
testit::assert(names(df) == c("species_tree", "gamma_stat"))
testit::assert(nrow(df) == 2)
knitr::kable(df)
```

## Collecting gamma's from the posterior

```{r}
filename <- find_path("toy_example_3.RDa")
df <- collect_posterior_gammas(filename)
testit::assert(names(df) == 
  c("species_tree", "alignment", "beast_run", "gamma_stat")
)
testit::assert(nrow(df) == 80)
knitr::kable(head(df))
```


## Collecting gamma's from a file

```{r}
filename <- find_path("toy_example_3.RDa")
df <- collect_file_gammas(filename)
testit::assert(names(df) == c("species_tree_gammas", "posterior_gammas"))
testit::assert(
  names(df$species_tree_gammas) == c("species_tree", "gamma_stat")
)
testit::assert(names(df$posterior_gammas) == 
  c("species_tree", "alignment", "beast_run", "gamma_stat")
)
testit::assert(nrow(df$species_tree_gammas) == 2)
testit::assert(nrow(df$posterior_gammas) == 80)
```

## Collecting gamma's from some files

```{r}
folder <- "/home/p230198/Peregrine"
all_parameter_filenames <- paste(folder, list.files(folder, pattern = "\\.RDa"), sep = "/")
df <- collect_files_gammas(head(all_parameter_filenames), verbose = TRUE)
testit::assert(names(df) 
  == c("species_tree_gamma_stats", "posterior_gamma_stats")
)
knitr::kable(df$species_tree_gamma_stats)
knitr::kable(head(df$posterior_gamma_stats))
```

## Collecting gamma's from all files

```{r}
csv_filename_species_trees <- "collected_gammas_species_trees.csv"
csv_filename_posterior <- "collected_gammas_posterior.csv"
zip_filename_posterior <- "collected_gammas_posterior.csv.zip"
create_fresh <- FALSE # Set this to TRUE and for a >60 mins calculation
if (create_fresh) {
  folder <- "/home/p230198/Peregrine"
  all_parameter_filenames <- paste(
    folder, list.files(folder, pattern = "\\.RDa"), sep = "/"
  )
  df <- collect_files_gammas(all_parameter_filenames, verbose = TRUE)
  write.csv(
    x = df$species_tree_gamma_stats,
    file = csv_filename_species_trees,
    row.names = TRUE
  )
  write.csv(
    x = df$posterior_gamma_stats,
    file = csv_filename_posterior,
    row.names = TRUE
  )
}
if (!file.exists(csv_filename_posterior)) {
  utils::unzip(zip_filename_posterior)
}
testit::assert(file.exists(csv_filename_species_trees))
testit::assert(file.exists(csv_filename_posterior))
```

Note that the comma-seperated file is only created when absent,
as this is a long step.

Loading it from file:

```{r}
df_species_trees <- read.csv(
 file = csv_filename_species_trees, 
 header = TRUE, 
 stringsAsFactors = FALSE, 
 row.names = 1
)

df_posterior <- read.csv(
 file = csv_filename_posterior, 
 header = TRUE, 
 stringsAsFactors = FALSE, 
 row.names = 1
)
```

Putting the head of the data frame in a table:

```{r}
knitr::kable(head(df_species_trees))
```

```{r}
knitr::kable(head(df_posterior))
```

What kind of data is `df_species_trees`?

```{r}
str(df_species_trees)
names(df_species_trees)
```

What kind of data is `df_posterior`?

```{r}
str(df_posterior)
names(df_posterior)
```

Let's first plot the distribution of gamma statistics.

```{r}
ggplot2::ggplot(
  data = df_species_trees, ggplot2::aes(df_species_trees$gamma_stat)
) + ggplot2::geom_histogram(binwidth = 1)
```

```{r}
ggplot2::ggplot(
  data = df_posterior, ggplot2::aes(df_posterior$gamma_stat)
) + ggplot2::geom_histogram(binwidth = 1)
```

## Richel's thoughts

I would be interested in the species trees with a gamma statistic
less than `-5`:

```{r}
species_trees_with_low_gamma <- df_species_trees[ 
  !is.na(df_species_trees$gamma_stat) & df_species_trees$gamma_stat < -5, 
]
knitr::kable(head(species_trees_with_low_gamma))
```

Which parameters go with that?

First read all parameters:

```{r}
csv_filename_parameters <- "collected_parameters.csv"
testit::assert(file.exists(csv_filename_parameters))
df_parameters <- read.csv(
 file = csv_filename_parameters, 
 header = TRUE, 
 stringsAsFactors = FALSE, 
 row.names = 1
)
knitr::kable(head(df_parameters))
```

Show the parameters that caused this:

```{r}
parameters_with_low_gamma <- df_parameters[ 
  rownames(df_parameters) %in% species_trees_with_low_gamma$filename,
]
knitr::kable(head(parameters_with_low_gamma))
```