---
title: "Analyse gamma's"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyse gamma's}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Loading this package

```{r}
library(Cer2016)
library(ggplot2)
options(warn = 2)
```

```{r}
csv_filename_species_trees <- find_path("collected_gammas_species_trees.csv")
csv_filename_posterior     <- find_path("collected_gammas_posterior.csv")
testit::assert(file.exists(csv_filename_species_trees))
testit::assert(file.exists(csv_filename_posterior))
```

Note that the comma-seperated file is only created when absent,
as this is a long step.

Loading it from file:

```{r}
df_species_trees <- read.csv(
 file = csv_filename_species_trees, 
 header = TRUE, 
 stringsAsFactors = FALSE, 
 row.names = 1
)

df_posterior <- read.csv(
 file = csv_filename_posterior, 
 header = TRUE, 
 stringsAsFactors = FALSE, 
 row.names = 1
)

df_parameters <- read_collected_parameters()

df_comparison <- merge(
  x = df_species_trees, 
  y = df_posterior, 
  by = c("filename", "species_tree"), 
  all = TRUE
)
df_comparison$diff <- df_comparison$gamma_stat.x - 
  df_comparison$gamma_stat.y
```

Putting the head of the data frame in a table:

```{r}
knitr::kable(head(df_species_trees))
```

```{r}
knitr::kable(head(df_posterior))
```

What kind of data is `df_species_trees`?

```{r}
str(df_species_trees)
names(df_species_trees)
```

What kind of data is `df_posterior`?

```{r}
str(df_posterior)
names(df_posterior)
```

Let's first plot the distribution of gamma statistics.

```{r fig.width = 7, fig.height = 7}
ggplot2::ggplot(
  data = df_species_trees, ggplot2::aes(df_species_trees$gamma_stat)) +   
  ggplot2::xlim(-15, 5) +
  ggplot2::geom_histogram(binwidth = 1,
                          colour = "grey38", fill = "gold2") +
  ggplot2::theme(text = ggplot2::element_text(size = 12)) +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 12)) +
  ggplot2::xlab("Gamma Statistic") +
  ggplot2::theme(axis.title.x = ggplot2::element_text(color="black",
                                             vjust=-0.35)) +
  ggplot2::ylab("Runs") +
  ggplot2::theme(axis.title.y = ggplot2::element_text(color="black" , 
                                             vjust=0.35)) +
  ggplot2::ggtitle("Sampled Trees Gamma Statistics") +
  ggplot2::theme(plot.title = ggplot2::element_text(size=12, 
                                  face="bold", 
                                  vjust=1, 
                                  lineheight=1))
```

```{r fig.width = 7, fig.height = 7}
ggplot2::ggplot(
  data = df_posterior, ggplot2::aes(df_posterior$gamma_stat)) + 
  ggplot2::xlim(-15, 5) +
  ggplot2::geom_histogram(binwidth = 1, 
                          colour = "grey38", 
                          fill = "gold2") + 
  ggplot2::theme(text = ggplot2::element_text(size = 12)) +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 12)) +
  ggplot2::xlab("Gamma Statistic") +
  ggplot2::theme(axis.title.x = ggplot2::element_text(color="black",
                                             vjust=-0.35)) +
  ggplot2::ylab("Runs") +
  ggplot2::theme(axis.title.y = ggplot2::element_text(color="black" , 
                                             vjust=0.35)) +
  ggplot2::ggtitle("Posterior Trees Gamma Statistics") +
  ggplot2::theme(plot.title = ggplot2::element_text(size=12, 
                                  face="bold", 
                                  vjust=1, 
                                  lineheight=1))
```

Hey, some of those gamma statistics are quite different between the species 
trees and the posterior! Let's take a look at that.

```{r}
counter <- 0
interesting_values <- NULL
for (stat in df_species_trees$gamma_stat){
  counter <- counter + 1
  if ( !is.na(stat) && (stat < -10)){
    interesting_values$filename  <- c(interesting_values$filename,
                                       df_species_trees$filename[counter])
    interesting_values$gamma_stat <- c(interesting_values$gamma_stat, stat)
    interesting_values$treenumber <- c(interesting_values$treenumber,
                                       df_species_trees$species_tree[counter])
  }
}
interesting_values <- data.frame(interesting_values)
```

```{r}
knitr::kable(head(interesting_values))
```

It seems like there are clusters of files that have very low gamma statistics. 
Let's compare them to the posterior values for those statistics.

```{r}
counter    <- 0
outliers   <- NULL
for (posterior in df_posterior$filename){
  counter  <- counter + 1
  counter2 <- 0
  for (filename in interesting_values$filename){
    counter2 <- counter2 + 1
    if (posterior == filename){
      outliers$filename  <- c(outliers$filename,
                               df_posterior$filename[counter])
      outliers$gamma_post <- c(outliers$gamma_post,
                               df_posterior$gamma_stat[counter])
      outliers$gamma_int  <- c(outliers$gamma_int,
                               interesting_values$gamma_stat[counter2])
      outliers$treenumber <- c(outliers$treenumber,
                               interesting_values$treenumber[counter2])
    }
  }
}

outliers <- data.frame(outliers)

print(names(outliers))
```

```{r}
knitr::kable(head(outliers))
```

All the low outliers don't seem to have a posterior... so the difference in the
graphs might not actually mean anything. Let's take a different approach. Let's
compare the gamma statistics of every tree that has a posterior.

```{r}
comparison  <- NULL
for (stat in df_posterior$gamma_stat) {
  counter <- 0
  if (!is.na(stat)){ 
    for (value in head(df_species_trees$gamma_stat)){
      counter <- counter + 1
      if (!is.na(value)){
        comparison$filename  <- c(comparison$filename,
                                   df_species_trees$filename[counter])
        comparison$treenumber <- c(comparison$treenumber,
                                   df_species_trees$species_tree[counter])
        comparison$gamma_pbd  <- c(comparison$gamma_pbd, value)
        comparison$gamma_post <- c(comparison$gamma_post, stat)
        comparison$diff       <- c(comparison$diff, 
                                   (value - stat))
      }
    }
  }
}
comparison <- data.frame(comparison)
```

```{r}
knitr::kable(head(comparison))
```

We actually used a different script for this. At any rate, this is what you get
if you plot it:

```{r}
ggplot2::qplot(diff, data = df_comparison, binwidth = 0.005, xlim = c(-2, 2))
```

And this is what you get if you plot the BD portion:

```{r fig.width = 7, fig.height = 7}
BD <- df_parameters[
  !is.na(df_parameters$scr) 
  & df_parameters$scr > 2,
  ]

knitr::kable(head(BD))

BD_gammas <- df_comparison[
  df_comparison$filename %in% rownames(BD),
  ]

knitr::kable(head(BD_gammas))

ggplot2::qplot(diff, data = BD_gammas, binwidth = 0.005, xlim = c(-2, 2))
```

So it seems like this is just noise.

```{r fig.width = 7, fig.height = 7}
PBD <- df_parameters[
  !is.na(df_parameters$scr)
  & df_parameters$scr < 2 
  & df_parameters$scr > 0,
  ]

knitr::kable(head(PBD))

PBD_gammas <- df_comparison[
  df_comparison$filename %in% rownames(PBD),
  ]

knitr::kable(head(PBD_gammas))

ggplot2::qplot(diff, data = PBD_gammas, binwidth = 0.005, xlim = c(-2, 2))

```

Mmmh. Small difference, PBD has a little less on the right of 0.

The lower the 'comparison' value, the bigger the error of BEAST towards higher 
gamma statistics.

Interestingly, some parameter values don't show up in the lower comparison 
values. The higher values for the speciation initiation rate for example 
(0.8 and 1). This turned to be because of missing posterior gammas for those
parameters. All files after `article_1_5_5_2_2.RDA` appear to be corrupted.

```{r fig.width = 7, fig.height = 7}
gamma_bigdiff <- df_comparison[ 
  !is.na(df_comparison$diff) & df_comparison$diff < -1, 
  ]

parameters_bigdiff <- df_parameters[ 
  rownames(df_parameters) %in% gamma_bigdiff$filename,
  ]

ggplot2::qplot(sirg, data = parameters_bigdiff, 
               binwidth = 0.1, xlim = c(0, 2))
```

This turned to be because of missing posterior gammas for those
parameters. All files after ```article_1_5_5_2_2.RDA``` appear to be corrupted.

```{r}
existing_post <- df_comparison[
  !is.na(df_comparison$diff),
  ]
tail(existing_post)
```

But back to comparing the parameters of the worse predicted gammas to those of 
the better predicted gammas.

```{r}
gamma_smalldiff <- df_comparison[
  !is.na(df_comparison$diff) & df_comparison$diff > -1,
  ]

parameters_smalldiff <- df_parameters[
  rownames(df_parameters) %in% gamma_smalldiff$filename,
  ]
```

Most of the time, there doesn't seem to be a difference. 

But let's look at age: It seems like the longer ```age``` is less well 
predicted.

```{r fig.width = 7, fig.height = 7}
ggplot2::qplot(age, data = parameters_bigdiff,
               binwidth = 0.5)

ggplot2::qplot(age, data = parameters_smalldiff,
               binwidth = 0.5)
``` 

Higher mutation rates are predicted better.

```{r fig.width = 7, fig.height = 7}
ggplot2::qplot(mutation_rate, data = parameters_bigdiff,
               binwidth = 0.005)

ggplot2::qplot(mutation_rate, data = parameters_smalldiff,
               binwidth = 0.005)
```

But those are the only parameters that seem to havce any kind of pattern to 
them.

## Richel's thoughts

I would be interested in the species trees with a gamma statistic
less than `-5`:

```{r}
species_trees_with_low_gamma <- df_species_trees[ 
  !is.na(df_species_trees$gamma_stat) & df_species_trees$gamma_stat < -5, 
]
knitr::kable(head(species_trees_with_low_gamma))
```

Which parameters go with that?

First read all parameters:

```{r}
df_parameters <- read_collected_parameters()
knitr::kable(head(df_parameters))
```

Show the parameters that caused this:

```{r}
parameters_with_low_gamma <- df_parameters[ 
  rownames(df_parameters) %in% species_trees_with_low_gamma$filename,
]
knitr::kable(head(parameters_with_low_gamma))
```

Let's look at what parameters are often there in the corrupted files.

```{r}
na_sample    <- NULL
counter      <- 0
for (stat in df_species_trees$gamma_stat){
  counter <- counter + 1
  if (!is.na(stat)){
    counter2 <- 0
    for (filename in row.names(df_parameters)){
      counter2 <- counter2 + 1
      if (filename == df_species_trees$filename[counter]) {
        na_sample$filename <- c(na_sample$filename, df_species_trees$filename[counter]) #nolint
        na_sample$treenumber <- c(na_sample$treenumber, df_species_trees$species_tree[counter]) #nolint
        na_sample$gamma_stat <- c(na_sample$gamma_stat, stat)
        na_sample$SIR_GS <- c(na_sample$SIR_GS, df_parameters$sirg[counter2])
        na_sample$SIR_IS <- c(na_sample$SIR_IS, df_parameters$siri[counter2])
        na_sample$SCR <- c(na_sample$SCR, df_parameters$scr[counter2])
        na_sample$EXT_GS <- c(na_sample$EXT_GS, df_parameters$erg[counter2])
        na_sample$EXT_IS <- c(na_sample$EXT_IS, df_parameters$eri[counter2])
        na_sample$AGE <- c(na_sample$AGE, df_parameters$age[counter2])
        na_sample$MUTATIONR <- c(na_sample$MUTATIONR, df_parameters$mutation_rate[counter2]) #nolint
        break
      }
    }
  }
}
na_sample <- data.frame(na_sample)
```

```{r}
knitr::kable(head(na_sample))
```

It seems like there is one parameter which is unusually often present in corrupt
files: a species initiation rate of 0.4. For some reason, these are very often
corrupted. This is very unexpected, as we'd expect something going wrong with
the program if the computer would be overwhelmed by calculations - for example
if the speciation initiatoin rate is high. But high speciation rates seem to be
fine. We later found out that this is caused by most of the 0.4 being 
overwritten before the experiment,
but not all, and the double dose of super-duper high speciation 
completion time runs. In the end, it turns out that the corrupted files are 
random... we are working with an involuntarily sampled set from the 'real' data. 

```{r fig.width = 7, fig.height = 7}
ggplot2::ggplot(
  data = na_sample, ggplot2::aes(SIR_GS)) +
  ggplot2::xlim(-0.1, 1) +
  ggplot2::geom_histogram(binwidth = 0.05,
                          colour = "grey38", fill = "gold2") +
  ggplot2::theme(text = ggplot2::element_text(size = 12)) +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 12)) +
  ggplot2::xlab("SIR") +
  ggplot2::theme(axis.title.x = ggplot2::element_text(color="black",
                                             vjust=-0.35)) +
  ggplot2::ylab("Runs") +
  ggplot2::theme(axis.title.y = ggplot2::element_text(color="black" , 
                                             vjust=0.35)) +
  ggplot2::ggtitle("Number of runs for Speciation Initiation Rate (SIR)") +
  ggplot2::theme(plot.title = ggplot2::element_text(size=12, 
                                  face="bold", 
                                  vjust=1, 
                                  lineheight=1))
```

The speciation completion rate is also interesting.

```{r fig.width = 7, fig.height = 7}
 ggplot2::ggplot(
  data = na_sample, ggplot2::aes(SCR)) +
  ggplot2::xlim(c(999999, 1000001)) +
  ggplot2::geom_histogram(binwidth = 0.05,
                          colour = "grey38", fill = "gold2") +
  ggplot2::theme(text = ggplot2::element_text(size = 12)) +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 12)) +
  ggplot2::xlab("SCR") +
  ggplot2::theme(axis.title.x = ggplot2::element_text(color="black",
                                             vjust=-0.35)) +
  ggplot2::ylab("Runs") +
  ggplot2::theme(axis.title.y = ggplot2::element_text(color="black" , 
                                             vjust=0.35)) +
  ggplot2::ggtitle("Number of runs for Speciation Completion Rate (SCR)") +
  ggplot2::theme(plot.title = ggplot2::element_text(size=12, 
                                  face="bold", 
                                  vjust=1, 
                                  lineheight=1))

ggplot2::ggplot(
  data = na_sample, ggplot2::aes(SCR)) +
  ggplot2::xlim(c(-0.1, 2)) +
  ggplot2::geom_histogram(binwidth = 0.05,
                          colour = "grey38", fill = "gold2") +
  ggplot2::theme(text = ggplot2::element_text(size = 12)) +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 12)) +
  ggplot2::xlab("SCR") +
  ggplot2::theme(axis.title.x = ggplot2::element_text(color="black",
                                             vjust=-0.35)) +
  ggplot2::ylab("Runs") +
  ggplot2::theme(axis.title.y = ggplot2::element_text(color="black" , 
                                             vjust=0.35)) +
  ggplot2::ggtitle("Number of runs for Speciation Completion Rate (SCR)") +
  ggplot2::theme(plot.title = ggplot2::element_text(size=12, 
                                  face="bold", 
                                  vjust=1, 
                                  lineheight=1))
```

```{r}
BD           <- NULL
counter      <- 0
for (rate in df_parameters$scr){
  counter <- counter + 1
  if (!is.na(rate) && rate > 1){
    BD$files <- c(BD$files, rownames(df_parameters)[counter])
  }  
}

counter <- 0
for (file in df_species_trees$filename){
  counter <- counter + 1
  for (name in BD$files)
    if (!is.na(df_species_trees$gamma_stat[counter]) && file == name){
      BD$filename  <- c(BD$filename, file)
      BD$gammas     <- c(BD$gammas, df_species_trees$gamma_stat[counter])
    } 
}
BD$files <- NULL

BD <- data.frame(BD)
knitr::kable(head(BD))

PBD    <- NULL
counter      <- 0
for (rate in df_parameters$scr){
  counter <- counter + 1
  if (!is.na(rate) && rate == 0.1){
    PBD$files <- c(PBD$files, rownames(df_parameters)[counter])
  }
}

counter <- 0
for (file in df_species_trees$filename){
  counter <- counter + 1
  for (name in PBD$files)
    if (!is.na(df_species_trees$gamma_stat[counter]) && file == name){
      PBD$filename  <- c(PBD$filename, file)
      PBD$gammas     <- c(PBD$gammas, df_species_trees$gamma_stat[counter])
    }
}
PBD$files <- NULL

PBD <- data.frame(PBD)

knitr::kable(head(PBD))
```

```{r fig.width = 7, fig.height = 7}
ggplot2::ggplot(
  data = PBD, ggplot2::aes(gammas)) +
  ggplot2::xlim(c(-15, 3)) +
  ggplot2::geom_histogram(binwidth = 0.05,
                          colour = "grey38", fill = "gold2") +
  ggplot2::theme(text = ggplot2::element_text(size = 12)) +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 12)) +
  ggplot2::xlab("Gammas") +
  ggplot2::theme(axis.title.x = ggplot2::element_text(color="black",
                                             vjust=-0.35)) +
  ggplot2::ylab("Runs") +
  ggplot2::theme(axis.title.y = ggplot2::element_text(color="black" , 
                                             vjust=0.35)) +
  ggplot2::ggtitle("Distribution of PBD gammas") +
  ggplot2::theme(plot.title = ggplot2::element_text(size=12, 
                                  face="bold", 
                                  vjust=1, 
                                  lineheight=1))

ggplot2::ggplot(
  data = BD, ggplot2::aes(gammas)) +
  ggplot2::xlim(c(-15, 3)) +
  ggplot2::geom_histogram(binwidth = 0.05,
                          colour = "grey38", fill = "gold2") +
  ggplot2::theme(text = ggplot2::element_text(size = 12)) +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 12)) +
  ggplot2::xlab("Gammas") +
  ggplot2::theme(axis.title.x = ggplot2::element_text(color="black",
                                             vjust=-0.35)) +
  ggplot2::ylab("Runs") +
  ggplot2::theme(axis.title.y = ggplot2::element_text(color="black" , 
                                             vjust=0.35)) +
  ggplot2::ggtitle("Distribution of BD gammas") +
  ggplot2::theme(plot.title = ggplot2::element_text(size=12, 
                                  face="bold", 
                                  vjust=1, 
                                  lineheight=1))
```

```{r fig.width = 7, fig.height = 7}
ggplot2::qplot(gammas, data = PBD, binwidth = 0.05, xlim = c(-15, 3))
ggplot2::qplot(gammas, data = BD, binwidth = 0.05, xlim = c(-15, 3))

pbds_with_low_gamma <- PBD[ 
  !is.na(PBD$gammas) & PBD$gammas < -10, 
]
pbds_with_low_gamma

parameters_with_low_gamma <- df_parameters[ 
  rownames(df_parameters) %in% pbds_with_low_gamma$filename,
]
parameters_with_low_gamma
```

## Richel's notes

```{r}
df_parameters_richel <- df_parameters[
  !is.na(df_parameters$scr)
  & df_parameters$scr > 0,
  ]

knitr::kable(head(df_parameters_richel))

richel_gammas <- df_comparison[
  df_comparison$filename %in% rownames(df_parameters_richel),
  ]

knitr::kable(head(richel_gammas))

```

## More of Femke her analysis

Analysis here:

```{r}

# Add a column called 'filenames' to the parameter files
df_parameters$filename <- rownames(df_parameters)

df_big <- merge(x = df_parameters, y = df_comparison, by = "filename", all = TRUE)

names(df_big)
knitr::kable(head(df_big))
df_big$scr <- as.factor(df_big$scr)

df_01 <- df_big[
  !is.na(df_big$scr)
  & (df_big$scr == 0.1 |       
     df_big$scr == 1e+06),
  ]

df_02 <- df_big[
  !is.na(df_big$scr)
  & (df_big$scr == 0.2 |       
     df_big$scr == 1e+06),
  ]

df_03 <- df_big[
  !is.na(df_big$scr)
  & (df_big$scr == 0.3 |       
     df_big$scr == 1e+06),
  ]

df_04 <- df_big[
  !is.na(df_big$scr)
  & (df_big$scr == 0.4 |       
     df_big$scr == 1e+06),
  ]

df_08 <- df_big[
  !is.na(df_big$scr)
  & (df_big$scr == 0.8 |       
     df_big$scr == 1e+06),
  ]

df_10 <- df_big[
  !is.na(df_big$scr)
  & (df_big$scr == 1 |       
     df_big$scr == 1e+06),
  ]

df_final <- df_big[
  !is.na(df_big$scr)
  & (df_big$scr == 0.2 |
     df_big$scr == 0.4 |
     df_big$scr == 0.8 |
     df_big$scr == 1e+06),
  ]
```

```{r fig.width = 7, fig.height = 7}
ggplot(df_final, aes(x = diff, colour = scr)) + 
  geom_density() +
  xlim(c(-1, 0.2)) +
  ggtitle("Comparison PBD - BD") +
  xlab("Diff sampled tree - posterior")

ggplot(df_final, aes(x = diff, colour = scr)) + 
  geom_density() +
  xlim(c(-2, 2)) +
  ggtitle("Comparison PBD - BD") +
  xlab("Diff sampled tree - posterior")

ggplot(df_big, aes(x = diff, colour = scr)) + 
  geom_density() +
  xlim(c(-0.2, 0.2)) +
  ggtitle("Comparison PBD - BD") +
  xlab("Diff sampled tree - posterior")

ggplot(df_big, aes(x = diff, colour = scr)) + 
  geom_density() +
  xlim(c(-2, 2)) +
  ggtitle("Comparison gamma statistic PBD - BD") +
  xlab("Diff sampled tree - posterior")

```

