---
title: "Analyse nLTTs"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyse nLTTs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Loading this package

```{r}
library(Cer2016)
```

Resolution of analysis:

```{r}
dt <- 0.1
```

## Collecting nLTT values from one or more phylogenies

```{r}
phylogenies <- c(ape::rcoal(10), ape::rcoal(20))
df <- ribir::get_nltt_values(phylogenies, dt = dt)
knitr::kable(df)
```

## Collecting nLTT values from the species trees

```{r}
filename <- find_path("toy_example_3.RDa")
df <- collect_species_tree_nltts(filename, dt = dt)
testit::assert(names(df) == c("species_tree", "t", "nltt"))
testit::assert(nrow(df) == 2 * (1 + (1 / dt)))
knitr::kable(df)
```

## Collecting nLTT values from the posterior

```{r}
filename <- find_path("toy_example_3.RDa")
df <- collect_posterior_nltts(filename, dt = dt)
testit::assert(names(df) == 
  c("species_tree", "alignment", "beast_run", "state", "t", "nltt")
)
testit::assert(nrow(df) == 880)
knitr::kable(head(df))
```

## Collecting nLTTs's from a file

```{r}
filename <- find_path("toy_example_3.RDa")
df <- collect_file_nltts(filename, dt = dt)
testit::assert(names(df) == c("species_tree_nltts", "posterior_nltts"))
testit::assert(
  names(df$species_tree_nltts) == c("species_tree", "t", "nltt")
)
testit::assert(names(df$posterior_nltts) == 
  c("species_tree", "alignment", "beast_run", "state", "t", "nltt")
)
testit::assert(nrow(df$species_tree_nltts) > 0)
testit::assert(nrow(df$posterior_nltts) > 0)
knitr::kable(head(df$species_tree_nltts))
knitr::kable(head(df$posterior_nltts))
```

## Collecting nLTTs from some files

```{r}
folder <- "/home/p230198/Peregrine"
all_parameter_filenames <- paste(folder, list.files(folder, pattern = "\\.RDa"), sep = "/")
df <- collect_files_nltts(head(all_parameter_filenames), verbose = TRUE)
testit::assert(names(df) 
  == c("species_tree_nltts", "posterior_nltts")
)
knitr::kable(head(df$species_tree_nltts))
knitr::kable(head(df$posterior_nltts))
```

## Collecting gamma's from all files

```{r}
csv_filename_species_trees <- "collected_nltts_species_trees.csv"
csv_filename_posterior <- "collected_nltts_posterior.csv"
zip_filename_posterior <- "collected_nltts_posterior.csv.zip"
create_fresh <- FALSE # Set this to TRUE and for a >60 mins calculation
if (create_fresh) {
  folder <- "/home/p230198/Peregrine"
  all_parameter_filenames <- paste(
    folder, list.files(folder, pattern = "\\.RDa"), sep = "/"
  )
  df <- collect_files_gammas(all_parameter_filenames, verbose = TRUE)
  write.csv(
    x = df$species_tree_nltts,
    file = csv_filename_species_trees,
    row.names = TRUE
  )
  write.csv(
    x = df$posterior_nltts,
    file = csv_filename_posterior,
    row.names = TRUE
  )
}
if (!file.exists(csv_filename_posterior)) {
  utils::unzip(zip_filename_posterior)
}
#testit::assert(file.exists(csv_filename_species_trees))
#testit::assert(file.exists(csv_filename_posterior))
```

Note that the comma-seperated file is only created when absent,
as this is a long step.

Loading it from file:

```{r}
# df_species_trees <- read.csv(
#  file = csv_filename_species_trees, 
#  header = TRUE, 
#  stringsAsFactors = FALSE, 
#  row.names = 1
# )
# 
# df_posterior <- read.csv(
#  file = csv_filename_posterior, 
#  header = TRUE, 
#  stringsAsFactors = FALSE, 
#  row.names = 1
# )
```

Putting the head of the data frame in a table:

```{r}
# knitr::kable(head(df_species_trees))
```

```{r}
# knitr::kable(head(df_posterior))
```

What kind of data is `df_species_trees`?

```{r}
# str(df_species_trees)
# names(df_species_trees)
```

What kind of data is `df_posterior`?

```{r}
# str(df_posterior)
# names(df_posterior)
```

Let's first plot the distribution of gamma statistics.

```{r}
# ggplot2::ggplot(
#   data = df_species_trees, ggplot2::aes(df_species_trees$nltt)
# ) + ggplot2::geom_histogram(binwidth = 1)
```

```{r}
# ggplot2::ggplot(
#   data = df_posterior, ggplot2::aes(df_posterior$nltt)
# ) + ggplot2::geom_histogram(binwidth = 1)
```

