---
title: "Analyse nLTTs"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyse nLTTs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Abstract

This vignette shows how to work with nLTT values.
It does *not* show the nLTT statistic, as that is
shown in the vignette 'Analyse nLTT statistics'.

## Setup

Loading this package

```{r}
library(Cer2016)
options(warn = 2)
```

## Collecting nLTTs from all files

Loading it from file:

```{r}
df_species_trees <- read_collected_nltts_strees()
df_posterior <- read_collected_nltts_postrs()
```

Putting the head of the data frame in a table:

```{r}
knitr::kable(head(df_species_trees))
knitr::kable(tail(df_species_trees))
```

```{r}
knitr::kable(head(df_posterior))
knitr::kable(tail(df_posterior))
```

Here we can see the nLTTs of the species trees used in the analysis:

```{r fig.width = 7, fig.height = 7}

df_species_trees$species_tree <- plyr::revalue(df_species_trees$species_tree, c("1"="youngest", "2"="oldest"))

ggplot2::ggplot(
  data = df_species_trees,
  ggplot2::aes(
    x = df_species_trees$t,
    y = df_species_trees$nltt,
    colour = df_species_trees$filename,
    shape = df_species_trees$species_tree
  )
) + ggplot2::geom_point(
) + ggplot2::geom_step(direction = "vh", stat = "summary", fun.y = "mean"
) + ggplot2::scale_x_continuous(
  limits = c(0, 1), name = "Time (normalized)"
) + ggplot2::scale_y_continuous(
  limits = c(0, 1), name = "nLTT"
) + ggplot2::ggtitle("Collected species trees") + 
  ggplot2::guides(colour = FALSE)
```

For every parameter combination (in this case summarized as a filename), there are two lines: one for the youngest and
one for the oldest subspiecies to represent a species.

Here we do the same for posteriors:

```{r fig.width = 7, fig.height = 7}
df_posterior$species_tree <- plyr::revalue(df_posterior$species_tree, c("1"="youngest", "2"="oldest"))

# Only select those with non-NA values
df <- df_posterior[ !is.na(df_posterior$t), ]

ggplot2::ggplot(
  data = df,
  ggplot2::aes(
    x = df$t,
    y = df$nltt,
    color = df$species_tree
  )
) + ggplot2::geom_point(
) + ggplot2::geom_step(direction = "vh", stat = "summary", fun.y = "mean"
) + ggplot2::scale_x_continuous(
  limits = c(0, 1), name = "Time (normalized)"
) + ggplot2::scale_y_continuous(
  limits = c(0, 1), name = "nLTT"
) + ggplot2::ggtitle("Collected posteriors"
) + ggplot2::geom_smooth()
  
```

Note that the smoother is not perfectly correct, but
I keep it in to give an idea of the distribition.
