---
title: "Analyse the number of taxa"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyse the number of taxa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Loading this package

```{r}
library(Cer2016)
```

## Collecting the number of taxa from one or more phylogenies

```{r}
phylogenies <- c(ape::rcoal(10), ape::rcoal(20))
df <- collect_n_taxa(phylogenies)
testit::assert(names(df) == c("n_taxa"))
testit::assert(df == data.frame(n_taxa = c(10, 20)))
knitr::kable(df)
```

## Collecting the number of taxa from the species trees

Because all species trees have the same number
of taxa, there will be one `n_taxa` per file.

```{r}
filename <- find_path("toy_example_3.RDa")
df <- collect_species_tree_n_taxa(filename)
testit::assert(names(df) == c("n_taxa"))
testit::assert(nrow(df) == 1)
knitr::kable(df)
```

To be sure, the species tree is plotted here:

```{r}
file <- read_file(filename)
plot(file$species_trees_with_outgroup[[1]][[1]])
```

## Collecting the number of taxa from the posterior

Because all posteriors have the same number
of taxa as the species trees, this step does
not make sense.

To be sure, the first phylogeny in the first posterior is plotted here:

```{r}
trees_filename <- find_path("toy_example_3_1_1_1.trees")
testit::assert(file.exists(trees_filename))
all_trees <- rBEAST::beast2out.read.trees(trees_filename)
plot(all_trees[[1]])
```

## Collecting the number of taxa from a file

Because all posteriors have the same number
of taxa as the species trees, this step does
not make sense.

## Collecting the number of taxa from some files

```{r}
filenames <- c(
  find_path("toy_example_1.RDa"),
  find_path("toy_example_2.RDa"),
  find_path("toy_example_3.RDa"),
  find_path("toy_example_4.RDa")
)

df <- collect_files_n_taxa(filenames, verbose = FALSE)
testit::assert(names(df) == c("filenames", "n_taxa"))
names(df)
knitr::kable(df)
```

## Collecting the number of taxa from all files

```{r}
csv_filename_n_taxa <- "collected_n_taxa.csv"
csv_filename_parameters <- "collected_parameters.csv"
create_fresh <- FALSE # Set this to TRUE for a >2 mins calculation
if (create_fresh) {
  folder <- "/home/p230198/Peregrine"
  all_parameter_filenames <- paste(
    folder, list.files(folder, pattern = "\\.RDa"), sep = "/"
  )
  df <- collect_files_n_taxa(all_parameter_filenames, verbose = FALSE)
  write.csv(
    x = df,
    file = csv_filename_n_taxa,
    row.names = TRUE
  )
}
testit::assert(file.exists(csv_filename_n_taxa))
```

Note that the comma-seperated file is only created when absent,
as this is a long step.

Loading it from file:

```{r}
df_n_taxa <- read.csv(
 file = csv_filename_n_taxa, 
 header = TRUE, 
 stringsAsFactors = FALSE, 
 row.names = 1
)

df_parameters <- read.csv(
  file = csv_filename_parameters, 
  header = TRUE, 
  stringsAsFactors = FALSE, 
  row.names = 1
)
```

Putting the head of the data frame in a table:

```{r}
knitr::kable(head(df_n_taxa))
```

What kind of data is `df_n_taxa`?

```{r}
str(df_n_taxa)
names(df_n_taxa)
```

Let's first plot the distribution of number of taxa statistics.

```{r}
useful_n_taxa <- df_n_taxa[ !is.na(df_n_taxa$n_taxa), ]
ggplot2::ggplot(
  data = useful_n_taxa, ggplot2::aes(useful_n_taxa$n_taxa)
) + ggplot2::geom_histogram(bins = 30)
```

```{r}
ggplot2::ggplot(
  data = useful_n_taxa, ggplot2::aes(useful_n_taxa$n_taxa)
) + ggplot2::geom_density()
```
