---
title: "branch_sorter"
author: "Femke Thon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## loading required packages
```{r}
library(ape)
library(Cer2016)
```

## generating a tree
```{r}
tree1 <- rtree(3)
plot(tree1)
```

We've got a tree. It's even got some info about itself in a file, which we'll 
need later.

```{r}
names(tree1)
```

We see quite a bit of info here already. In fact, everything is already there.
By combining ```'edge.length'``` with ```'edge'``` and ```'tip.label'``` we will 
be able to sort and extract the info we need for the NRBS analysis.

```{r}
tree1.sorted <- branch_sorter(tree1)
```

We've got a sorted tree! How wonderful! What does it look like?

```{r}
names(tree1.sorted)

tree1.sorted$branch_length

tree1.sorted$taxon_name
```

There we have it. ```Tree1.sorted$branch_length``` gives us the lengths of the
branches, ```tree1.sorted$taxon_name``` the names that correspond to these
branches.

## comparing the tree to itself.
We want to compare it to itself, so we duplicate it. (once I get the NRBS to 
work again.)

```{r}
tree2 <- tree1
```

