---
title: "convert_alignment_to_beast_posterior demo"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{convert_alignment_to_beast_posterior demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

To convert an alignment to a BEAST2 posterior:

 * Create a phylogeny
 * Create an alignment from that phylogeny
 * Create a BEAST2 posterior from that alignment
 * Compare the BEAST2 posterior with the original phylogeny

## Load the packages

```{r}
library(Cer2016)
library(nLTT)
library(phangorn)
library(rBEAST)
library(testit)
```


## Create a phylogeny

```{r}
n_taxa <- 5
phylogeny <- ape::rcoal(n = n_taxa)
plot(phylogeny)
```

## Create an alignment from that phylogeny

```{r}
alignment <- convert_phylogeny_to_alignment(
  phylogeny = phylogeny,
  sequence_length = 10
)
image(alignment)
```

From that alignment, create a BEAST2 posterior

```{r}
base_filename <- "convert_alignment_to_beast_posterior_test"

if (file.exists(paste(base_filename, ".log", sep = ""))) {
  file.remove(paste(base_filename, ".log", sep = ""))  
}
if (file.exists(paste(base_filename, ".trees", sep = ""))) {
  file.remove(paste(base_filename, ".trees", sep = ""))  
}
if (file.exists(paste(base_filename, ".xml.state", sep = ""))) {
  file.remove(paste(base_filename, ".xml.state", sep = ""))  
}
posterior <- convert_alignment_to_beast_posterior(
  alignment,
  mcmc_chainlength = 10000,
  base_filename = base_filename,
  rng_seed = 42
)
```

## Plot this posterioir

```{r}
last_tree <- tail(posterior, n = 1)[[1]]
plot(last_tree, main = "Last tree in posterior")

all_nltt_stats <- NULL
for (tree in posterior)
{
  all_nltt_stats = c(all_nltt_stats, nLTT::nLTTstat(phylogeny, tree))
}

hist(all_nltt_stats)
```

## Cleanup

```{r}
if (file.exists(paste(base_filename, ".log", sep = ""))) {
  file.remove(paste(base_filename, ".log", sep = ""))  
}
if (file.exists(paste(base_filename, ".trees", sep = ""))) {
  file.remove(paste(base_filename, ".trees", sep = ""))  
}
if (file.exists(paste(base_filename, ".xml.state", sep = ""))) {
  file.remove(paste(base_filename, ".xml.state", sep = ""))  
}
```