---
title: "Demo of the summary statistics"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demo of the summary statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

There are multiple phylogeny summary statistics:

 * [demo], by Richel
 * [first], by Femke
 * [second], by Jolien
 
## Richel's demo summary statistic

First, we read a parameter file:

```{r}
library(Cer2016)
filename <- find_path("toy_example_1.RDa")
file <- read_file(filename)
```

A file has some properties:

```{r}
print(names(file))
```

In this example, I compare the nLTT of the unsampled original species
tree (independent if this is a good idea) to the average nLTT
of the posterior.

The unsampled original species tree is in `file$pbd_output`, let's
see what's there:

```{r}
print(names(file$pbd_output))
```

I want to get the incipient species tree of the extant species

```
file$pbd_output$igtree.extant
```



## [First summary statistic]

Given a phylogeny [TODO: Use a BD or PBD phylogeny]

```{r}
library(ape)
phylogeny <- rcoal(10)
plot(phylogeny)
```

A possible summary statistic is ...

The statistic is very in-/sensitive, as when we compare the phylogeny
to itself, the values already differ????

```{r}
measurement_1 <- runif(n = 1, max = sum(phylogeny$edge.length))
measurement_2 <- runif(n = 1, max = sum(phylogeny$edge.length))
print(measurement_1)
print(measurement_2)
```






