---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette exists to describe and fix problem. It fixes
problems just by using `Knitr` on it :-)

## `Cannot find package beastscriptr`

Install it like this:

```{r}
if (is.element("beastscriptr", installed.packages()[, 1])) {
  print("package is already installed")
} else {
  devtools::install_github("richelbilderbeek/beastscriptr")
}
```

## `Cannot find package rBEAST`

Perhaps the prerequisites are missing:

```{r}
if (is.element("adephylo", installed.packages()[, 1])) {
  print("package is already installed")
} else {
  install.packages("adephylo", repos = "http://cran.uk.r-project.org")
}
```

Then install it like this:

```{r}
if (is.element("rBEAST", installed.packages()[, 1])) {
  print("package is already installed")
} else {
  devtools::install_github("richelbilderbeek/rBEAST")
}
```


## `Cannot find package ribir`

Install it like this:

```{r}
if (is.element("ribir", installed.packages()[, 1])) {
  print("package is already installed")
} else {
  devtools::install_github("richelbilderbeek/ribir")
}
```

## `Cannot find xml2-config`

This may happen when you install the `XML` package:

```
install.packages("XML")
[...]
checking for xml2-config... no
Cannot find xml2-config
```

Solution (from [Stack Overflow](http://stackoverflow.com/questions/7765429/unable-to-install-r-package-in-ubuntu-11-04)) 
is to install `libxml2-dev` on the command line.

```
sudo apt-get install libxml2-dev
```

Note: an alternative is to use

```
sudo apt-get install r-cran-xml
```
